Literature Review

### 2018-06-07 LiteratureReview/LiteratureReview3000.tex
made changes to literature review

\title{Literature Review}
\date{2018-04-26}
\author{Jomar Alcantara}

@@

\par
According to a recent report commissioned by the Alzheimer's Society, they estimate the prevalence of Dementia in the UK at approximately 815,000 people. This represents 1 in 14 of the population aged 65 or over. This report also estimates an annual healthcare spend on £4.3 billion of which approximately £85 million is spend solely on diagnosis. They also estimate that the overall impact of dementia (excluding the costs associated with early onset dementia) is £26.3 billion annually. Globally, this picture is a lot bleaker. A recent report suggests that in 2015 there were 46 million people with a diagnosis of dementia and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}. The report also states that the worldwide cost of dementia in 2018 is estimated to be in the region of one trillon US dollars.
\newline
\par
The purpose of this review is to seek to understand how dementia impacts speech and language production from a psychological and linguistic perspective with a view to potentially using natural language processing and technology to aid in the early diagnosis of dementia and/or Mild Cognitive Impairment.
\newline
\par
A search of the literature was conducted mainly from a psychological perspective as this would inform the theoretical side of things. Next, a search of the literature from a Computer Science perspective was also made with a view to exploring the how domain of machine learning and more specifically natural language processing can be applied to this problem. Finally a search was carried out to see what work has already been done in this area.

@@

\section{Description of Dementia}

Language - DSM-V
'Major' Has significant difficulties with expressive or receptive language. Often uses general-use phrases such as "that thing" and "you know what I mean," and prefers general pronouns rather than names. With severe impairment, may not even recall names of closer friends and family. Idiosyncratic word usage, grammatical errors, and spontaneity of output and economy of utterances occur. Stereotypy of speech occurs; echolalia and automatic speech typically precede mutism.

'Mild' Has noticeable word-finding difficulty. May substitute general for specific terms. May avoid use of specific names of acquaintances. Grammatical errors involve subtle omission or incorrect use of articles, prepositions, auxiliary verbs, etc.

\subsection{Alzheimer's Disease}
\subsection{Vascular Dementia}
\subsection{Dementia with Lewy Bodies}
\subsection{Parkinson's disease dementia}
\subsection{Fronto-temporal dementia}
\subsection{Primary Progressive Aphasia}
\subsection{Mild Cognitive Impairment}

@@

Emery \cite{Emery2000}, states that you can see language decline as Hierarchical. According to a semiotic model there are four ranks of language based on their cognitive complexity. These ranks are Phonology, Morphology, Syntax and Semantics. She conducted a review studies which looked at language decline from the perspective of these hierarchical ranks. Her conclusion is that language decline is related to the complexity of the language task given to a participant and that language decline is hierarchical in that the language forms we learn last (the most complex language forms) are the first to deteriorate.

A common way for clinicians' to diagnose dementia is through use of the Mini-Mental State Examination. The Mini-Mental State Examination (MMSE) has shown to be particularly effective at differentiating betweeen Dementia and other psychological disorders such as clinical depression, schitzophrenia and personality disorder and is currently the most widely used measure of diagnosing Dementia within clinical psychology.  However, it is not without it's faults...

\section{Current ways of diagnosing Dementia using Language}
\subsection{Free Cued Selective Recall Test}
The Free Cued Selective Recall Test (FCSRT) was created by Grober and his team \cite{Grober1988}. This is a test that primarily tests memory but also tests a subjects ability to name objects. Four items are presented at a time on a card and each participant is asked to name and recall with and without cues.

\subsection{Mini-Mental State Examination}

Battery of tests to differentiate between Controls, those with MCI, those with AD. 
\begin{itemize}
	\item FCSRDT - As most effective 
	\item Mini Mental State Examination (MMSE)
	\item Addenbrooke's Cognitive Examination (ACE-III)
	\item Dementia Rating Scale (DRS)
	\item Western Aphasia Battery (WAB)
	\item Boston Diagnostic Aphasia Examination (BDAE)
	\item Boston Naming Test (BNT)
\end{itemize}

\section{Existing studies that have used Natural Language Processing to help in diagnosing psychological difficulties, particularly language difficulties}

\subsection{Nun Study - Snowdon (1996)}
\subsection{Comparing Presidents Speeches - Berisha et al (2015)}

\section{State of literature into Natural Language processing techniques}

\section{stuff to add}
Annotated Bibliography for the project "Ambient Monitoring of Speech for Evidence of Cognitive Decline".

Orimaye, S. O., Wong, J. S-M., Golden, K. J., Wong, C. P., & Soyiri, I. N. (2017) Predicting probably Alzheimer's Disease using linguistic deficits and biomarkers.

@@

Asgari, M., Kaye, J., & Dodge, H. (2017) Predicting mild cognitive impairment from spontaneous spoken utterances. Alzheimers & Dementia: Translational Research & Clinical Interventions 3, 219 - 228

Asgari, Kaye and Dodge (2017) also looked at the linguistic characteristics of older adults with mild cognitive impairment (MCI) vs healthy controls where they hypothesised that they would be able to predict those with MCI, a distinguishing characteristic of Alzheimer's disease and other variations of dementia. Using recordings of unstructured conversations (with standardized preselected topics across subjects) between interviewers and interviewees they grouped spoken words using Linguistic Inquiry and Word Count (LIWC) which is a technique used to categorize words into features such as negative and positive words. They then applied support vector machines (SVM's) and random forest classification algorithms to investigate whether machine learning could be used to distinguish between those with MCI and healthy controls. They were able to successfully used machine learning algorithms to distinguish between these two groups with an accuracy of 84%. The authors report that this method is highly reliant on high-fidelity transcription of the conversations which is labour intensive, but they anticipate that technology is advancing sufficiently quickly that automated high-quality transcriptions are possible in the near future. This paper provides a different perspective in how to tackle the problem deriving language features from speech using a different framework. The challenges both Orimaye et al and these authors experienced were around the automation of transcription and this would potentially be an area to explore in the proposed project.

Pennebaker, J.W., Boyd, R. L., Jordan, K and Blackburn, K., (2015) The development and psychometric properties of LIWC2015

Asgari et al (2017) used the Linguistic Inquiry and Word Count (LIWC) methodology, developed by Pennebaker et al (2007) to generate features for their paper. Pennebaker, Boyd, Jordan and Blackburn (2015) have since developed their framework. The LIWC is designed to categorize and evaluate the various emotional, cognitive and structural components which are present within samples of speech and/or written text. This method, Linquistic Inquiry & Word Count (LIWC) has evolved from it's initial incarnation in 1993 to it's last update in 2015. The premise is the use of words from particular categories provide an insight into the psychological processes and/or diagnoses an individual has. The latest version of this methodology uses a dictionary of 6400 words, word stems and select emoticons and assigns these constructs to various categories. This has been updated to include modern uses of language such as 'text speak'. This manual goes summarises the process of constructing the framework as well as discussing the reliability and validity of the measure and a review of studies using the LIWC was conducted by Tausczik and Pennebaker (2010) support the notion that the LIWC is valid across multiple psychological domains. This paper is useful as it provides a useful and validated framework from which the proposed project can potentially derive language features.

Schwartz, A. H., Eichstaedt, J., Kern, M. L., Park, G., Sap. M., Stillwell. D., . . . Ungar. L. (2014)  Towards Assessing Changes in Degree of Depression through Facebook. Workshop on Computational Linguistics and Clinical Psychology: From Linguistic Signal to Clinical Reality, 118-125 
Currently formal diagnosis of depression is expensive and time-consuming involving both the use of questionnaires and the use of a trained professional to assess an individual. Schwartz et al, explored the use of language as an aid to diagnose depression in a naturalistic setting i.e. facebook status updates. The authors used both the LIWC (described above) and n-grams as features to distinguish those with depression from healthy individuals and were able to track participants levels' of depression through language successfully.Whilst the authors were able to use a reasonably sound method of measure degree of depression and have evidenced a methodology that is able to track a levels of depression over time, it would be more helpful to use a more clinically relevant measure of depression and anxiety i.e., PHQ-9 (Kroenke, 2001) and GAD-7 (Spitzer et al, 2006) to add validity to their findings. 
This paper is interesting as it uses both LIWC and n-grams as a method of deriving features from text / speech for depression which, as described above, have also been used as features to distinguish MCI from healthy participants. Given the comorbity of depression and dementia (Meyers, 1998), it would be interesting for the proposed project to further analyse the link between dementia and clinical depression via the use of language. This paper also supports the idea that language changes over time and is a marker for deterioration in both those with depression and dementia.

@@

\bibliographystyle{unsrt}
\bibliography{LiteratureReview}
\end {document}

>>>>

\title{Literature Review}
\date{2018-06-06}
\author{Jomar Alcantara}

@@

\par
According to a recent report commissioned by the Alzheimer's Society, they estimate the prevalence of Dementia in the UK at approximately 815,000 people. This represents 1 in 14 of the population aged 65 or over. This report also estimates an annual healthcare spend on £4.3 billion of which approximately £85 million is spend solely on diagnosis. They also estimate that the overall impact of dementia (excluding the costs associated with early onset dementia) is £26.3 billion annually. Globally, this picture is a lot bleaker. A recent report suggests that in 2015 there were 46 million people with a diagnosis of dementia and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}. The report also states that the worldwide cost of dementia in 2018 is estimated to be in the region of one trillon US dollars. \newline

According to the DSM 5 \cite{AmericanPsychiatricAssociation2013}, those with Mild dementia suffer from noticeable word-find difficulty. They may substitute general terms for more specific terms. They may avoid the use of specific names of acquaintances. There may be grammatical errors involving subtle omission or incorrect use of articles, prepositions, auxiliary verbs, etc. Those who have progressed from Mild to Major depression also have difficulties with expressive or receptive language. They will often use general-use phrases such as "that thing" and "you know what I mean," and prefers general pronouns rather than names. With severe impairment, sufferers may not even recall names of closer friends and family. Idiosyncratic word usage, grammatical errors, and spontaneity of output and economy of utterances occur. Stereotypy of speech occurs; echolalia and automatic speech typically precede mutism.
\newline
\par
The purpose of this review is to seek to understand how dementia impacts speech and language production from a psychological and linguistic perspective. A search of the literature was conducted mainly from a psychological perspective using ProQuest (PsychArticles), SCOPUS, Web of Science. The following results were found.

\begin{center}
	\begin{tabular}{ | l | l | l | p{3cm} |}
		\hline
		Database & Number of Results & Search Terms  \\ \hline
		ProQuest(PsychArticles) & 1484 Results & Language AND Decline AND Dementia \\ \hline
		ProQuest(PsychArticles) & 486 Results  & Language AND Decline AND Dementia AND Speech \\ \hline
		Web of Science & 1207 Results  & Language AND Decline AND Dementia   \\ \hline
		Web of Science & 151 Results  & Language AND Decline AND Dementia AND Speech  \\ \hline
		Scopus & 791 Results & Language AND Decline AND Dementia  \\ \hline
		Scopus & 91 Results & Language AND Decline AND Dementia AND Speech   \\ 
		\hline
	\end{tabular}
\end{center}

@@

@@

Emery (2000)  \cite{Emery2000}, takes the perspective of semiotics, and the idea that language and the processes involved in language are hierarchical. She proposes that language goes from simple units of construction, and build layers of complexity and sophistication. There are four levels of language according to this perspective: Phonology, Morphology, Syntax and Semantics. Emery conducted a review of research that looked at how AD impacted on each level of language. She found that people with AD generally had intact Phonology and Morphology but more impaired Syntax and Semantics.

Emery states that you can see language decline as Hierarchical. According to a semiotic model there are four ranks of language based on their cognitive complexity. These ranks are Phonology, Morphology, Syntax and Semantics. She conducted a review studies which looked at language decline from the perspective of these hierarchical ranks. Her conclusion is that language decline is related to the complexity of the language task given to a participant and that language decline is hierarchical in that the language forms we learn last (the most complex language forms) are the first to deteriorate.

\section{Connected Speech}

\section{stuff to add}
Annotated Bibliography for the project "Ambient Monitoring of Speech for Evidence of Cognitive Decline".
\newline
\par 

@@

\newline
\par
Asgari, Kaye and Dodge (2017) also looked at the linguistic characteristics of older adults with mild cognitive impairment (MCI) vs healthy controls where they hypothesised that they would be able to predict those with MCI, a distinguishing characteristic of Alzheimer's disease and other variations of dementia. Using recordings of unstructured conversations (with standardized preselected topics across subjects) between interviewers and interviewees they grouped spoken words using Linguistic Inquiry and Word Count (LIWC) which is a technique used to categorize words into features such as negative and positive words. They then applied support vector machines (SVM's) and random forest classification algorithms to investigate whether machine learning could be used to distinguish between those with MCI and healthy controls. They were able to successfully used machine learning algorithms to distinguish between these two groups with an accuracy of 84. The authors report that this method is highly reliant on high-fidelity transcription of the conversations which is labour intensive, but they anticipate that technology is advancing sufficiently quickly that automated high-quality transcriptions are possible in the near future. This paper provides a different perspective in how to tackle the problem deriving language features from speech using a different framework. The challenges both Orimaye et al and these authors experienced were around the automation of transcription and this would potentially be an area to explore in the proposed project. Asgari et al (2017) used the Linguistic Inquiry and Word Count (LIWC) methodology, developed by Pennebaker et al (2007) to generate features for their paper. Pennebaker, Boyd, Jordan and Blackburn (2015) have since developed their framework. The LIWC is designed to categorize and evaluate the various emotional, cognitive and structural components which are present within samples of speech and/or written text. This method, Linquistic Inquiry and Word Count (LIWC) has evolved from it's initial incarnation in 1993 to it's last update in 2015. The premise is the use of words from particular categories provide an insight into the psychological processes and/or diagnoses an individual has. The latest version of this methodology uses a dictionary of 6400 words, word stems and select emoticons and assigns these constructs to various categories. This has been updated to include modern uses of language such as 'text speak'. This manual goes summarises the process of constructing the framework as well as discussing the reliability and validity of the measure and a review of studies using the LIWC was conducted by Tausczik and Pennebaker (2010) support the notion that the LIWC is valid across multiple psychological domains. This paper is useful as it provides a useful and validated framework from which the proposed project can potentially derive language features.
\newline
\par
Currently formal diagnosis of depression is expensive and time-consuming involving both the use of questionnaires and the use of a trained professional to assess an individual. Schwartz et al, explored the use of language as an aid to diagnose depression in a naturalistic setting i.e. facebook status updates. The authors used both the LIWC (described above) and n-grams as features to distinguish those with depression from healthy individuals and were able to track participants levels' of depression through language successfully.Whilst the authors were able to use a reasonably sound method of measure degree of depression and have evidenced a methodology that is able to track a levels of depression over time, it would be more helpful to use a more clinically relevant measure of depression and anxiety i.e., PHQ-9 (Kroenke, 2001) and GAD-7 (Spitzer et al, 2006) to add validity to their findings. 
This paper is interesting as it uses both LIWC and n-grams as a method of deriving features from text / speech for depression which, as described above, have also been used as features to distinguish MCI from healthy participants. Given the comorbity of depression and dementia (Meyers, 1998), it would be interesting for the proposed project to further analyse the link between dementia and clinical depression via the use of language. This paper also supports the idea that language changes over time and is a marker for deterioration in both those with depression and dementia.

@@

\bibliographystyle{unsrt}
\bibliography{LiteratureReview3000}
\end {document}
